Gammapy: A Python package for gamma-ray astronomy

### 2023-01-27 src/ms-review.tex
Add Appendix with code example output

%% CN: Shouldn't we display the output of this snippet?
%% otherwise what's the point of the `print`' functions?
\begin{figure}
	\small
	\import{code-examples/generated/}{gp_data}
	\caption{
        Using \code{gammapy.data} to access DL3 level data with a \code{DataStore} object.
        Individual observations can be accessed by their unique integer observation id number.
        The actual events and instrument response functions can be accessed
        as attributes on the \code{Observation} object, such as \code{.events}
        or \code{.aeff} for the effective area information.
    }
	\label{fig*:minted:gp_data}
\end{figure}
%

@@

\subsubsection{Effective Area}
\gammapy provides the class \code{EffectiveAreaTable2D} to
manage the effective area, which is usually defined in terms of true energy and offset angle.
The class functionalities offer the possibility to read from files or to create
it from scratch. {The \code{EffectiveAreaTable2D} class can also convert, interpolate,
write, and evaluate the effective area for a given energy and offset angles, or
even plot the multi-dimensional effective area table.

@@

To handle the spatial dimension the sub-package exposes a uniform API for
the FITS World Coordinate System (WCS), the HEALPix pixelization and
region-based data structure (see Figure~\ref{ig*:minted:gp_maps}).
This allows uses to perform the same higher level operations on maps
independent of the underlying pixelisation scheme.

@@

	\caption{
        Using \code{gammapy.maps} to create a WCS, a HEALPix and a region
		based data structures. The initialisation parameters include
        consistently the positions of the center of the map, the pixel
        size, the extend of the map as well as the energy axis definition.
        The energy minimum and maximum values for the creation of the
        \code{MapAxis} object can be defined as strings also specifying the
        unit. Region definitions can be passed as strings following
        the DS9 region specifications \url{http://ds9.si.edu/doc/ref/region.html}.
        }
    \label{ig*:minted:gp_maps}
\end{figure}

@@

\subsection{gammapy.datasets}
\label{ssec:gammapy-datasets}
%
\begin{figure}
	\small
	\import{code-examples/generated/}{gp_datasets}
	\caption{
        Using \code{gammapy.datasets} to read existing reduced binned datasets.
        After the different datasets are read from disk they are collected into a
        common \code{Datasets} container. All dataset types have an associated
        name attribute to allow access by name later in the code. The
        environment variable \code{\$GAMMAPY\_DATA} is automtically resolved
        by \gammapy.
    }
	\label{fig*:minted:gp_datasets}
\end{figure}
%
The \code{gammapy.datasets} subpackage contains classes to bundle
together binned data along with the associated models and likelihood function, which
provides an interface to the \code{Fit} class (Sec \ref{sssec:fit}) for
modeling and fitting purposes. Depending upon the type of analysis and the
associated statistic, different types of Datasets are supported. The \code{MapDataset} is
used for combined spectral and morphological (3D) fitting, while a 1D spectral 
fitting can be performed using the \code{SpectrumDataset}.
While the default fit statistics for both of these classes is the \emph{Cash}}~\citep{Cash}
statistic, there are other classes which support
analyses where the background is measured from control regions, so called "off" obervations.
Those require the use of a different fit statistics, which takes into account the
uncertainty of the background measurement. This case is covered by the \code{MapDatasetOnOff}
and \code{SpectrumDatasetOnOff} classes, which use the \emph{WStat}~\citep{WStat} statistic.

@@

\subsection{gammapy.makers}
\label{ssec:gammapy-makers}
%
\begin{figure}
	\small
    \import{code-examples/generated/}{gp_makers}
	\caption{
        Using \code{gammapy.makers} to reduce DL3 level data into a
		\code{MapDataset}. All \code{Maker} classes represent 
		a step in the data reduction process. They take
        the configuration on initialisation of the class. They 
		also consistently define \code{.run()} methods, which take
		a dataset object and optionally an \code{Observation} 
		object. In this way, \code{Maker} classes can be chained
		to define more complex data reduction pipelines.
    }
	\label{ig*:minted:gp_makers}
\end{figure}
%
The \code{gammapy.makers} sub-package contains the various classes and functions required
to process and prepare \gammaray data from the DL3 to the DL4, representing the input for modeling and fitting.
First, events are binned and IRFs are interpolated and projected onto the chosen analysis
geometry. The end product of the data reduction process are a set of binned counts,
background exposure, psf and energy dispersion maps at the DL4 level. 
The \code{MapDatasetMaker} and \code{SpectrumDatasetMaker} are
responsible for this task for 3D and 1D analyses, respectively (see Fig~\ref{ig*:minted:gp_makers}).

@@

\begin{figure}
	\small
	\import{code-examples/generated/}{gp_stats}
	\caption{
        Using \code{gammapy.stats} to compute statistical quantities
        such as excess, signficance and assymetric errors
        from counts based data. The data is passed on initialisation
        of the \code{WStatCountsStatistic} class. The quantities
        are the computed ON excess of the corresponding class
        attributes such as \code{stat.n\_sig} and \code{stat.sqrt\_ts}.
    }
	\label{codeexample:stats} \end{figure}
%% CN: the figure does not explain what alpha is

@@

%% CN: here again, print() without output. Is it useful?
\begin{figure}
	\small
	\import{code-examples/generated/}{gp_models}
	\caption{Using \code{gammapy.modeling.models} to define a source model with a
    spectral, spatial and temporal component. For convenience the model
    parameters can be defined as strings with attached units. The spatial model
    takes an additional \code{frame} parameter which allow users to define
    the coordinate frame of the position of the model.
    }
	\label{fig*:minted:gp_models}
\end{figure}

@@

\begin{figure}
	\small
	\import{code-examples/generated/}{gp_estimators}
	\caption{Using the \code{TSMapEstimator} object from \code{gammapy.estimators} to compute a
        a flux, flux upper limits and TS map. The additional parameters \code{n\_sigma}
        and \code{n\_sigma\_ul} define the confidence levels (in multiples of the normal distribution width)
        of the flux error and and flux upper limit maps respectively.
    }
    \label{fig*:minted:gp_estimators}
\end{figure}

@@

\begin{figure}
	\small
	\import{code-examples/generated/}{gp_catalogs}
	\caption{Using \code{gammapy.catalogs} to access the underlying model, flux points and
		light-curve from the \fermi 4FGL catalog for the blazar PKS 2155-304}
	\label{codeexample:data}
\end{figure}

@@

\appendix

>>>>

%% CN: Shouldn't we display the output of this snippet?
%% otherwise what's the point of the `print`' functions?
\begin{figure}
	\small
	\import{code-examples/generated/}{gp_data}
	\caption{
        Using \code{gammapy.data} to access DL3 level data with a \code{DataStore} object.
        Individual observations can be accessed by their unique integer observation id number.
        The actual events and instrument response functions can be accessed
        as attributes on the \code{Observation} object, such as \code{.events}
        or \code{.aeff} for the effective area information. The output
		of the code example is shown in Figure~\ref{fig:code_example_gp_data}.
    }
	\label{fig*:minted:gp_data}
\end{figure}
%

@@

\subsubsection{Effective Area}
\gammapy provides the class \code{EffectiveAreaTable2D} to
manage the effective area, which is usually defined in terms of true energy and offset angle.
The class functionalities offer the possibility to read from files or to create
it from scratch. The \code{EffectiveAreaTable2D} class can also convert, interpolate,
write, and evaluate the effective area for a given energy and offset angles, or
even plot the multi-dimensional effective area table.

@@

To handle the spatial dimension the sub-package exposes a uniform API for
the FITS World Coordinate System (WCS), the HEALPix pixelization and
region-based data structure (see Figure~\ref{fig*:minted:gp_maps}).
This allows uses to perform the same higher level operations on maps
independent of the underlying pixelisation scheme.

@@

	\caption{
        Using \code{gammapy.maps} to create a WCS, a HEALPix and a region
		based data structures. The initialisation parameters include
        consistently the positions of the center of the map, the pixel
        size, the extend of the map as well as the energy axis definition.
        The energy minimum and maximum values for the creation of the
        \code{MapAxis} object can be defined as strings also specifying the
        unit. Region definitions can be passed as strings following
        the DS9 region specifications \url{http://ds9.si.edu/doc/ref/region.html}.The output
		of the code example is shown in Figure~\ref{fig:code_example_gp_maps}.
        }
    \label{fig*:minted:gp_maps}
\end{figure}

@@

\subsection{gammapy.datasets}
\label{ssec:gammapy-datasets}
%
\begin{figure}
	\small
	\import{code-examples/generated/}{gp_datasets}
	\caption{
        Using \code{gammapy.datasets} to read existing reduced binned datasets.
        After the different datasets are read from disk they are collected into a
        common \code{Datasets} container. All dataset types have an associated
        name attribute to allow access by name later in the code. The
        environment variable \code{\$GAMMAPY\_DATA} is automtically resolved
        by \gammapy. The output
		of the code example is shown in Figure~\ref{fig:code_example_gp_datasets}.
    }
	\label{fig*:minted:gp_datasets}
\end{figure}
%
The \code{gammapy.datasets} subpackage contains classes to bundle
together binned data along with the associated models and likelihood function, which
provides an interface to the \code{Fit} class (Sec \ref{sssec:fit}) for
modeling and fitting purposes. Depending upon the type of analysis and the
associated statistic, different types of Datasets are supported. The \code{MapDataset} is
used for combined spectral and morphological (3D) fitting, while a 1D spectral 
fitting can be performed using the \code{SpectrumDataset}.
While the default fit statistics for both of these classes is the \emph{Cash}~\citep{Cash}
statistic, there are other classes which support
analyses where the background is measured from control regions, so called "off" obervations.
Those require the use of a different fit statistics, which takes into account the
uncertainty of the background measurement. This case is covered by the \code{MapDatasetOnOff}
and \code{SpectrumDatasetOnOff} classes, which use the \emph{WStat}~\citep{WStat} statistic.

@@

\subsection{gammapy.makers}
\label{ssec:gammapy-makers}
%
\begin{figure}
	\small
    \import{code-examples/generated/}{gp_makers}
	\caption{
        Using \code{gammapy.makers} to reduce DL3 level data into a
		\code{MapDataset}. All \code{Maker} classes represent 
		a step in the data reduction process. They take
        the configuration on initialisation of the class. They 
		also consistently define \code{.run()} methods, which take
		a dataset object and optionally an \code{Observation} 
		object. In this way, \code{Maker} classes can be chained
		to define more complex data reduction pipelines. The output
		of the code example is shown in Figure~\ref{fig:code_example_gp_makers}.
    }
	\label{fig*:minted:gp_makers}
\end{figure}
%
The \code{gammapy.makers} sub-package contains the various classes and functions required
to process and prepare \gammaray data from the DL3 to the DL4, representing the input for modeling and fitting.
First, events are binned and IRFs are interpolated and projected onto the chosen analysis
geometry. The end product of the data reduction process are a set of binned counts,
background exposure, psf and energy dispersion maps at the DL4 level. 
The \code{MapDatasetMaker} and \code{SpectrumDatasetMaker} are
responsible for this task for 3D and 1D analyses, respectively (see Figure~\ref{fig*:minted:gp_makers}).

@@

\begin{figure}
	\small
	\import{code-examples/generated/}{gp_stats}
	\caption{
        Using \code{gammapy.stats} to compute statistical quantities
        such as excess, signficance and assymetric errors
        from counts based data. The data is passed on initialisation
        of the \code{WStatCountsStatistic} class. The quantities
        are the computed ON excess of the corresponding class
        attributes such as \code{stat.n\_sig} and \code{stat.sqrt\_ts}.
		The output of the code example is shown in Figure~\ref{fig:code_example_gp_stats}.
    }
	\label{fig*:minted:gp_stats}
\end{figure}
%% CN: the figure does not explain what alpha is

@@

%% CN: here again, print() without output. Is it useful?
\begin{figure}
	\small
	\import{code-examples/generated/}{gp_models}
	\caption{Using \code{gammapy.modeling.models} to define a source model with a
    spectral, spatial and temporal component. For convenience the model
    parameters can be defined as strings with attached units. The spatial model
    takes an additional \code{frame} parameter which allow users to define
    the coordinate frame of the position of the model. The output
	of the code example is shown in Figure~\ref{fig:code_example_gp_models}.
    }
	\label{fig*:minted:gp_models}
\end{figure}

@@

\begin{figure}
	\small
	\import{code-examples/generated/}{gp_estimators}
	\caption{Using the \code{TSMapEstimator} object from \code{gammapy.estimators} to compute a
        a flux, flux upper limits and TS map. The additional parameters \code{n\_sigma}
        and \code{n\_sigma\_ul} define the confidence levels (in multiples of the normal distribution width)
        of the flux error and and flux upper limit maps respectively. The output
		of the code example is shown in Figure~\ref{fig:code_example_gp_estimators}.
    }
    \label{fig*:minted:gp_estimators}
\end{figure}

@@

\begin{figure}
	\small
	\import{code-examples/generated/}{gp_catalogs}
	\caption{Using \code{gammapy.catalogs} to access the underlying model, flux points and
		light-curve from the \fermi 4FGL catalog for the blazar PKS 2155-304. The output
		of the code example is shown in Figure~\ref{fig:code_example_gp_catalogs}.
	}
	\label{fig*:minted:gp_catalogs}
\end{figure}

@@

\appendix
\section{Code Examples Output}

\begin{figure}[!ht]
	\small
	\import{code-examples/generated-output/}{gp_data}
	\caption{Output from the code example shown in Figure~\ref{fig*:minted:gp_data}}
	\label{fig:code_example_gp_data}
\end{figure}

\begin{figure}[!ht]
	\small
	\import{code-examples/generated-output/}{gp_datasets}
	\caption{Output from the code example shown in Figure~\ref{fig*:minted:gp_datasets}}
	\label{fig:code_example_gp_datasets}
\end{figure}

\begin{figure}[!ht]
	\small
	\import{code-examples/generated-output/}{gp_maps}
	\caption{Output from the code example shown in Figure~\ref{fig*:minted:gp_maps}}
	\label{fig:code_example_gp_maps}
\end{figure}

\begin{figure}[!ht]
	\small
	\import{code-examples/generated-output/}{gp_stats}
	\caption{Output from the code example shown in Figure~\ref{fig*:minted:gp_stats}}
	\label{fig:code_example_gp_stats}
\end{figure}

\begin{figure*}[!ht]
	\centering
	\includegraphics[width=1.\textwidth]{figures/gp_makers.pdf}
	\caption{Output from the code example shown in Figure~\ref{fig*:minted:gp_makers}}
	\label{fig:code_example_gp_makers}
\end{figure*}

\begin{figure*}[!ht]
	\centering
	\includegraphics[width=1.\textwidth]{figures/gp_estimators.pdf}
	\caption{Output from the code example shown in Figure~\ref{fig*:minted:gp_estimators}}
	\label{fig:code_example_gp_estimators}
\end{figure*}

\begin{figure*}[!ht]
	\centering
	\includegraphics[width=1.\textwidth]{figures/gp_catalogs.pdf}
	\caption{Output from the code example shown in Figure~\ref{fig*:minted:gp_catalogs}}
	\label{fig:code_example_gp_catalogs}
\end{figure*}

\begin{figure}[!ht]
	\small
	\import{code-examples/generated-output/}{gp_models}
	\caption{Output from the code example shown in Figure~\ref{fig*:minted:gp_models}}
	\label{fig:code_example_gp_models}
\end{figure}
